---
title: "Code Used to create descriptive tables in the COVID DSL v3 manuscript"
author: Miguel Ángel Armengol de la Hoz / Jesús Villar Fernández
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

## Libraries

```{r message=FALSE, warning=FALSE}
library(summarytools)
library(RPostgreSQL)
library(tidyverse)
library(stringr)
library(caret)
library(tableone)
library(hrbrthemes)
library(plotly)
library(readxl)
library(xlsx)
```

## Database connection

This notebook requires the database and schema available in the github [COVID HM Hospitales Code Repository](https://github.com/theonesp/covidhm-code/) to be be deployed to be able to run the following code.

### Drivers

```{r}
#install.packages("RPostgreSQL")
library(RPostgreSQL)

# Load configuration settings
dbdriver <- 'PostgreSQL'
host  <- 'localhost'
port  <- '45432'
user  <- 'postgres'
password <- 'Baba0805'
dbname <- 'covidhm'
schema <- 'covidhm_dsl_v3' #schema <- 'covid_dsl_v2'
encoding  <- 'latin1'

# Connect to the database using the configuration settings
con <- dbConnect(dbDriver(dbdriver), dbname = dbname, host = host, port = port, 
                 user = user, password = password)
rm(password) # removes the password


# Set the default schema
dbExecute(con, paste("SET search_path TO ", schema, sep=" "))

# WARNING: accents presents in the table need special encoding

dbGetQuery(con, "SET client_encoding = 'LATIN1'")

dbGetQuery(con, "SHOW CLIENT_ENCODING")
```

## Creating useful query functions

```{r}
# function avoiding add the connection data in each iteration
run_query<-function(query){
query_output<-dbGetQuery(con,query)
  return(query_output)
}

# function for reading sql files
getSQL <- function(filepath){
  con = file(filepath, "r")
  sql.string <- ""

  while (TRUE){
    line <- readLines(con, n = 1)

    if ( length(line) == 0 ){
      break
    }

    line <- gsub("\\t", " ", line)

    if(grepl("--",line) == TRUE){
      line <- paste(sub("--","/*",line),"*/")
    }

    sql.string <- paste(sql.string, line)
  }

  close(con)
  return(sql.string)
}
```




# Table 1

Patient characteristics.

## Data extraction

```{r}
#patient<-run_query(getSQL('../sql/patient_v3.sql'))
patient<-run_query(getSQL('../sql/patient_survived_collapsed_v3.sql'))
```

## Table 1 generation

Demographic table

```{r}
library(tableone)

contvars<-c("age","admission_d_inpat","icu_date_in","icu_date_out","icu_days","discharge_date","admission_date_emerg","department_emerg","diag_emerg","temp_first_emerg","hr_first_emerg","glu_first_emerg","sat_02_first_emerg","bp_max_first_emerg","bp_min_first_emerg","hr_last_emerg","temp_last_emerg","glu_last_emerg","sat_02_last_emerg","bp_max_last_emerg","bp_min_last_emerg")

catvars<-c("sex","diag_inpat","mechvent")

allvars<-c(contvars,catvars)

statifyby<-'destin_discharge'

table1 <- CreateTableOne(vars = allvars, data = patient, factorVars = catvars)

table1strat <- CreateTableOne(vars = allvars, data = patient, factorVars = catvars
                         , testNormal=oneway.test 
                         , strata = statifyby
                         , smd = TRUE
                         )

table1_printed<-as.data.frame(print(table1
                                    , nonnormal = c('age',"temp_first_emerg","hr_first_emerg","glu_first_emerg","sat_02_first_emerg","bp_max_first_emerg","bp_min_first_emerg","hr_last_emerg","temp_last_emerg","glu_last_emerg","sat_02#_last_emerg","bp_max_last_emerg","bp_min_last_emerg","icu_days")
                                    ,missing = T
                                    ,minMax = F))

table1strat_printed<-as.data.frame(print(table1strat
                                    , nonnormal = c('age',"temp_first_emerg","hr_first_emerg","glu_first_emerg","sat_02_first_emerg","bp_max_first_emerg","bp_min_first_emerg","hr_last_emerg","temp_last_emerg","glu_last_emerg","sat_02#_last_emerg","bp_max_last_emerg","bp_min_last_emerg","icu_days")
                                    ,missing = F
                                    ,minMax = F))

table1_final<-cbind(table1_printed,table1strat_printed)
table1_final$Missing_N<-round((as.numeric(table1_final['n',]$Overall)*as.numeric(table1_final$Missing))/100,0)

table1_final<-table1_final%>%rename(Missing_pct=Missing)

#table1_final
write.xlsx(as.data.frame(print(table1_final)), paste("covid_dsl_table1_final_", Sys.Date(),".xlsx",sep=''))
```

# Table 2: ICD

## Data extraction

```{r}
#hosp_adm_icd<-run_query(getSQL('../sql/hosp_adm_icd_v3.sql'))
#er_adm_icd<-run_query(getSQL('../sql/er_adm_icd_v3.sql'))

hosp_adm_icd<-run_query(getSQL('../sql/hosp_adm_icd_survived_collapsed_v3.sql'))
er_adm_icd<-run_query(getSQL('../sql/er_adm_icd_survived_collapsed_v3.sql'))
```

## Selecting top ICD

```{r}
top_hosp_adm_icd<-as.data.frame(table(hosp_adm_icd$abbrev_desc))
top_hosp_adm_icd<-top_hosp_adm_icd[order(top_hosp_adm_icd$Freq,decreasing = T),]
print(top_hosp_adm_icd)

# selecting only top5, pull converts dataframe into array
top_hosp_adm_icd<-pull(top_hosp_adm_icd %>% top_n(wt =Freq,n = 5)%>%select(Var1))
```

```{r}
top_er_adm_icd<-as.data.frame(table(er_adm_icd$abbrev_desc))
top_er_adm_icd<-top_er_adm_icd[order(top_er_adm_icd$Freq,decreasing = T),]
print(top_er_adm_icd)

# selecting only top5, pull converts dataframe into array
top_er_adm_icd<-pull(top_er_adm_icd %>% top_n(wt =Freq,n = 5)%>%select(Var1)) 
```

## Table 2A generation

ICD hosp table

```{r}
contvars<-c("")

catvars<-c("abbrev_desc")

allvars<-c(contvars,catvars)

statifyby<-'destin_discharge'

table1 <- CreateTableOne(vars = allvars
                         , data = hosp_adm_icd%>%filter(abbrev_desc %in% top_hosp_adm_icd)
                         , factorVars = catvars
                         )

table1strat <- CreateTableOne(
                           vars = allvars
                         , data = hosp_adm_icd%>%filter(abbrev_desc %in% top_hosp_adm_icd)
                         , factorVars = catvars
                         , testNormal=oneway.test 
                         , strata = statifyby
                         , smd = TRUE
                         )
table1_printed<-as.data.frame(print(table1
                                    ,missing = T
                                    ,minMax = F

                                    ))
table1strat_printed<-as.data.frame(print(table1strat
                                    ,missing = F
                                    ,minMax = F))

table1_final<-cbind(table1_printed,table1strat_printed)
table1_final$Missing_N<-round((as.numeric(table1_final['n',]$Overall)*as.numeric(table1_final$Missing))/100,0)

table1_final<-table1_final%>%rename(Missing_pct=Missing)

table1_final
write.xlsx(as.data.frame(print(table1_final)), paste("covid_dsl_table2a_final_", Sys.Date(),".xlsx",sep=''))
```

## Table 2B generation

ICD hosp table

```{r}
contvars<-c("")

catvars<-c("abbrev_desc")

allvars<-c(contvars,catvars)

statifyby<-'destin_discharge'

table1 <- CreateTableOne(vars = allvars
                         , data = er_adm_icd%>%filter(abbrev_desc %in% top_er_adm_icd)
                         , factorVars = catvars
                         ) 

table1strat <- CreateTableOne(vars = allvars
                         , data = er_adm_icd%>%filter(abbrev_desc %in% top_er_adm_icd)
                         , factorVars = catvars
                         , testNormal=oneway.test 
                         , strata = statifyby
                         , smd = TRUE
                         )

table1_printed<-as.data.frame(print(table1
                                    ,missing = T
                                    ,minMax = F
                                    ))
table1strat_printed<-as.data.frame(print(table1strat
                                    ,missing = F
                                    ,minMax = F))

table1_final<-cbind(table1_printed,table1strat_printed)
table1_final$Missing_N<-round((as.numeric(table1_final['n',]$Overall)*as.numeric(table1_final$Missing))/100,0)

table1_final<-table1_final%>%rename(Missing_pct=Missing)

table1_final
write.xlsx(as.data.frame(print(table1_final)), paste("covid_dsl_table2b_final_", Sys.Date(),".xlsx",sep=''))
```

# Table 3 Comorbidities

## Data extraction

```{r}
comorbidities_hosp_adm<-run_query(getSQL('../sql/comorbidities_hosp_adm_v3.sql'))
```

## Table 3 generation

Demographic table

```{r}
contvars<-c()

catvars<-c("pmhx_diabetes","pmhx_hld","pmhx_htn","pmhx_ihd","pmhx_ckd","pmhx_copd","pmhx_asthma","pmhx_activecancer","pmhx_chronicliver","pmhx_stroke")

allvars<-c(contvars,catvars)

#statifyby<-'destin_discharge'

table1 <- CreateTableOne(vars = allvars, data = comorbidities_hosp_adm, factorVars = catvars)
table1_printed<-as.data.frame(print(table1
                                    ,missing = T
                                    ,minMax = F
                                    ))

table1_printed$Missing_N<-round((as.numeric(table1_printed['n',]$Overall)*as.numeric(table1_printed$Missing))/100,0)

table1_printed<-table1_printed%>%rename(Missing_pct=Missing)

table1_printed
write.xlsx(as.data.frame(print(table1_printed)), paste("covid_dsl_table3_", Sys.Date(),".xlsx",sep=''))
```